### 2017-12-15 submission/manuscript.tex

\textbf{\emph{Statistical Analysis:}} All analysis was done with the R
(v 3.4.2) software package (R Core Team, 2017). Data tranformation and
graphing was completed using the tidyverse package (v 1.1.1) and colors
selected using the viridis package (v 0.4.0) (Garnier, 2017; Wickham,
2017). The total number of OTUs were analyzed using an ANOVA with a
tukey post-hoc test. For the fecal samples the data was normalized to
each individual by cycle number to account for the biological variation
between different people. Bray-Curtis distance matrices were generated
using mothur after 100 sub-samplings at 1000, 5000, 10000, and 15000.
The distance matrix data was analyzed using PERMANOVA with the vegan
package (v 2.4.4) (Oksanen et al., 2017) and kruskal-wallis tests within
R. For both error and chimera analysis, samples were tested using
Kruskal-Wallis with a Dunns post-hoc test. Where applicable correction
for multiple comparison utilized the Benjamini-Hochberg method
(Benjamini \& Hochberg, 1995).

\textbf{\emph{Analysis Workflow:}} The total number of OTUs after
sub-sampling was analyzed for both the fecal and mock community samples.
Cycle dependent affects on Bray-Curits distances were next assessed for
the fecal samples looking at both overall cycle differences and within
individual differences for the previous cycle (e.g.~20 versus 25, 25
versus 30, etc.). From these observations we wanted to next analyze
potential reasons as to why some of these differences may have occured.
First, analysis of general sequence error rate, number of sequences with
an error, and base subsitution were assessed in the mock community for
each Taq. After assessing these errors, the total number of chimeras was
determined after sequence processing. The fecal samples were analyzed at
4 different sub-sampling levels, 1000, 5000, 10000, and 15000 while the
mock community samples were analysed at 3 levels, 1000, 5000, 10000.

@@

\textbf{\emph{Minimal Bray-Curtis Differences are Detected by Cycle
Number:}} Using PERMANOVA to test whether there were any differences
within HiFi DNA polymerase groups based on cycle number, only Phusion
had cycle dependent differences at 1000 and 5000 sub-sampling depth
(P-value = 0.03 and 0.01, respectively). Interestingly, Phusion was one
of only two DNA polymerase that managed to have samples for the 1000
sub-sampling depth at 15 cycles. Next, we assessed whether there were
any major differences between 5 cycle increments within each sample. We
found that there was no detectable difference in distance when comparing
to the previous 5 cycle increment (P-value \textgreater{} 0.05).
However, Phusion at 1000 sub-sampling depth had a P-value = 0.02 before
multiple comparison correction. It should be noted that at higher
sub-sampling depths these differences in Bray-Curits distance disappear
for both differences in cycle number and within 5 cycle increments
within an individual.

@@

@@

>>>>

\textbf{\emph{Statistical Analysis:}} All analysis was done with the R
(v 3.4.2) software package (R Core Team, 2017). Data tranformation and
graphing was completed using the tidyverse package (v 1.1.1) and colors
selected using the viridis package (v 0.4.0) (Garnier, 2017; Wickham,
2017). The total number of OTUs were analyzed using an ANOVA with a
tukey post-hoc test. For the fecal samples the data was normalized to
each individual by cycle number to account for the biological variation
between different people. Bray-Curtis matrices were generated using
mothur after 100 sub-samplings at 1000, 5000, 10000, and 15000. The
distance matrix data was analyzed using PERMANOVA with the vegan package
(v 2.4.4) (Oksanen et al., 2017) and kruskal-wallis tests within R. For
both error and chimera analysis, samples were tested using
Kruskal-Wallis with a Dunns post-hoc test. Where applicable correction
for multiple comparison utilized the Benjamini-Hochberg method
(Benjamini \& Hochberg, 1995).

\textbf{\emph{Analysis Workflow:}} The total number of OTUs after
sub-sampling was analyzed for both the fecal and mock community samples.
Cycle dependent affects on Bray-Curits indicies were next assessed for
the fecal samples looking at both overall cycle differences and within
individual differences for the previous cycle (e.g.~20 versus 25, 25
versus 30, etc.). From these observations we wanted to next analyze
potential reasons as to why some of these differences may have occured.
First, analysis of general sequence error rate, number of sequences with
an error, and base subsitution were assessed in the mock community for
each Taq. After assessing these errors, the total number of chimeras was
determined after sequence processing. The fecal samples were analyzed at
4 different sub-sampling levels, 1000, 5000, 10000, and 15000 while the
mock community samples were analysed at 3 levels, 1000, 5000, 10000.

@@

\textbf{\emph{Minimal Bray-Curtis Differences are Detected by Cycle
Number:}} Using PERMANOVA to test whether there were any differences
within HiFi DNA polymerase groups based on cycle number, only Phusion
had cycle dependent differences at 1000 and 5000 sub-sampling depth
(P-value = 0.03 and 0.01, respectively). Interestingly, Phusion was one
of only two DNA polymerase that managed to have samples for the 1000
sub-sampling depth at 15 cycles. Next, we assessed whether there were
any major differences between 5 cycle increments within each sample. We
found that there was no detectable difference in Bray-Curits index when
comparing to the previous 5 cycle increment (P-value \textgreater{}
0.05). However, Phusion at 1000 sub-sampling depth had a P-value = 0.02
before multiple comparison correction. It should be noted that at higher
sub-sampling depths these differences in Bray-Curits indicies disappear
for both differences in cycle number and within 5 cycle increments
within an individual.

@@

Although we did not observe strong differences, based on cycle number,
using the Bray-Curtis index the data suggests that there may be
differences between 15 cycles and higher cycle numbers, such as 30, that
are commonly used. What is most interesting is that there was no
difference within individuals between corresponding 5 cycle increments
(e.g.~15 to 20, 20 to 25, etc.). This finding, in conjunction with the
PERMANOVA results, would suggest that differences based on Bray-Curtis
incrementaly build up so that within 5 cycle increments the differences
are not that large but when comparing larger cycle differences, as with
the PERMANOVA, these differences start to become detectable. Although
our data suggests that it is possible to overcome these types of cycle
dependent differences with increased sub-sampling depth, for some DNA
polymerase increased sub-sampling depth increases chimera prevalence and
per base error rate. Conversely, it is possible that many of the
increased number of OTUs, generated as cycle number increases, are not
highly abundant and the Bray-Curtis index is able to successfully
downweight these respective OTUs (Minchin, 1987).

@@

\hypertarget{ref-bc_index_Minchin1987}{}
Minchin PR. 1987. An evaluation of the relative robustness of techniques
for ecological ordination. \emph{Vegetatio} 69:89--107. DOI:
\href{https://doi.org/10.1007/bf00038690}{10.1007/bf00038690}.
